---
title: "Chapter 3 Questions"
author: "Melissa Van Bussel"
date: "June 4, 2018"
output:
  pdf_document:
    highlight: zenburn
    latex_engine: lualatex
  html_document: default
  word_document: default
mainfont: Arial
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Chapter 3 points: 2 + 7 + 9 = 18 
 
#\textcolor{cyan}{Easy (2 points total)}

The following code is given in the question: 

```{r}
library(rethinking)
p_grid <- seq(from = 0, to = 1, length.out = 1000) 
prior <- rep(1, 1000) 
likelihood <- dbinom(6 , size = 9, prob = p_grid) 
posterior <- likelihood * prior 
posterior <- posterior / sum(posterior) 
set.seed(100) 
samples <- sample(p_grid, prob = posterior, size = 1e4, replace = TRUE) 
```

##\textcolor{cyan}{3E1-3E7}

```{r}
question3e1 <- length(which(posterior < 0.2)) / length(posterior)
question3e2 <- length(which(posterior > 0.8)) / length(posterior)
question3e3 <- length(which(posterior > 0.2 & posterior < 0.8)) / length(posterior)
question3e4 <- quantile(posterior, 0.2)
question3e5 <- quantile(posterior, 0.8)

# Narrowest interval would be centred at 0.5, so we want 33% above and below 0.5
question3e6upper <- quantile(posterior, 0.5+0.33)
question3e6lower <- quantile(posterior, 0.5-0.33)

# Equal on both sides would mean there's (0.33/2) on either side. 
question3e7upper <- quantile(posterior, 1-(0.33/2))
question3e7lower <- quantile(posterior, 0.33/2)
```

#\textcolor{cyan}{Medium (9 points total)}

##\textcolor{cyan}{3M1 (2 points)}

```{r}
p_grid <- seq(from = 0, to = 1, length.out = 1000)
prior <- rep(1, 1000)
likelihood <- dbinom(8, size = 15, prob = p_grid)
unstd.posterior <- likelihood * prior 
posterior <- unstd.posterior / sum(unstd.posterior)
```

##\textcolor{cyan}{3M2 (2 points)}

```{r}
samples <- sample(x = p_grid, size = 10000, prob = posterior, replace = TRUE)
HPDI(samples, 0.9)
```

##\textcolor{cyan}{3M3 (1 point)}

```{r}
posterior_predictive_check <- rbinom(n = 10000, size = 15, prob = samples)
sum(posterior_predictive_check == 8) / 10000
```

##\textcolor{cyan}{3M4 (2 points)}

```{r}
posterior_predictive_check <- rbinom(n = 10000, size = 9, prob = samples)
sum(posterior_predictive_check == 6) / 10000
```

##\textcolor{cyan}{3M5 (2 points)}

```{r}
prior <- c(rep(0, 500), rep(1, 500))
likelihood <- dbinom(8, size = 15, prob = p_grid)
unstd.posterior <- likelihood * prior 
posterior <- unstd.posterior / sum(unstd.posterior)

# 3M2
samples <- sample(x = p_grid, size = 10000, prob = posterior, replace = TRUE)
HPDI(samples, 0.9)
# 3M3
posterior_predictive_check <- rbinom(n = 10000, size = 15, prob = samples)
sum(posterior_predictive_check == 8) / 10000
# 3M4
posterior_predictive_check <- rbinom(n = 10000, size = 9, prob = samples)
sum(posterior_predictive_check == 6) / 10000
```

#\textcolor{cyan}{Hard (7 points total)}

The following code is used in all of the following questions.

```{r}
library(rethinking) 
data(homeworkch3)
```

##\textcolor{cyan}{3H1 (1 point)}

```{r}
num_births <- length(birth1) + length(birth2)
num_boys <- sum(birth1) + sum(birth2)
prior <- rep(1, 1000)
likelihood <- dbinom(x = num_boys, size = num_births, prob = p_grid)
unstd.posterior <- prior * likelihood
posterior <- unstd.posterior / sum(unstd.posterior)

p_grid[which.max(posterior)]
```

##\textcolor{cyan}{3H2 (2 points)}

```{r}
my_sample <- sample(p_grid, 10000, prob = posterior, replace = TRUE)
HPDI(my_sample, 0.5)
HPDI(my_sample, 0.89)
HPDI(my_sample, 0.97)
```

##\textcolor{cyan}{3H3 (2 points)}

```{r}
simulation <- rbinom(10000, size = 200, prob = my_sample)
dens(simulation)
# yes, it looks like 111 is right near the centre of the distribution.
```

##\textcolor{cyan}{3H4 (1 point)}

```{r}
new_simulation <- rbinom(10000, size = 100, prob = my_sample)
dens(new_simulation)
```

##\textcolor{cyan}{3H5 (1 point)}

```{r}
boys_born_after_girls <- birth2[birth1 == 0]
posterior_predictive_check <- rbinom(n = 10000, size = length(boys_born_after_girls), prob = length(boys_born_after_girls) / length(birth2))
dens(posterior_predictive_check)
```